---
title: "The Recipe for Success in the NBA"
subtitle: "due Friday, Nov 20 at 11:59pm"
author: "Late Round Sleepers: Alex Jackson, Evan Finley, Jack Griffin, Aliya Kamran"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

# 1: Introduction

We are investigating what makes NBA players successful in context of their team 
play. The reason for choosing this study is that are all NBA fans and eager to 
apply our new knowledge of data science to the game will allow us to have a 
better understanding for the game and make us more informed viewers. 
Furthermore, basketball statistics are fairly new and it will be interesting to 
discover if our perceived notions of certain players match the statistical data. 

In particular, we defined our investigation of "success" of NBA players and team
on the following hypothesis tests and questions:

1. Does the average offensive and defensive raptor for players differ between 
the regular season and post season?

2. Does the average raptor differ greatly from the average WAR between members 
of the All NBA teams?

3. How are the variables pace impact, raptor total and war total impacted by 
player type?

4. Do better defensive players have a better WAR in the playoffs and regular 
season?

5. Do players with offensive capabilities have greater impact on team 
possessions per 48 minutes as compared to players without offensive 
capabilities?

6. Does the team's average raptor correlate to the teams average winning
percentage?

7. Does the team's average war correlate to the teams average winning
percentage?


8. Which statistic, total raptor or total WAR, is better at indicicating player 
success? 

Sources:
https://fivethirtyeight.com/features/introducing-raptor-our-new-metric-for-the-modern-nba/


# 2: Data description

```{r environ}
library(tidyverse)
nba_players <- read.csv("https://projects.fivethirtyeight.com/nba-model/2020/latest_RAPTOR_by_team.csv")
```

We are exploring the NBA-Raptor data set. Raptor is a player efficiency rating 
system.This data set has tracked multiple team and player analysis over a six 
year period. Each observation is an NBA player with variables about their 
statistics. It has variables such as players points per game, team winning 
percentage, average team possessions per game, and many more. The player data
is stats from every NBA games since 2014. Raptor was made by Fivethirtyeight. 

The data set has 808 observatios and 23 variables with the investigation focusing on the following variables:

`player_name`:	Player name
`season`:	Season
`season_type`:	Regular season (RS) or playoff (PO)
`team`:	Basketball-Reference ID of team
`poss`:	Possessions played
`mp`:	Minutes played
`raptor_offense`:	Points above average per 100 possessions added by player on offense, using both box and on-off components
`raptor_defense`:	Points above average per 100 possessions added by player on defense, using both box and on-off components
`raptor_total`:	Points above average per 100 possessions added by player on both offense and defense, using both box and on-off components
`war_total`:	Wins Above Replacement between regular season and playoffs
`war_reg_season`:	Wins Above Replacement for regular season
`war_playoffs`:	Wins Above Replacement for playoffs
`pace_impact`:	Player impact on team possessions per 48 minutes

## 3: Glimpse of data

```{r glimpse}
glimpse(nba_players)
```

# 4: Methodology

The methodology section should include the variables used to address your 
research question, as well as any useful visualizations or summary statistics. 
As well, you should introduce and justify the statistical method(s) that you 
believe will be useful in answering your research question.


To start with, we looked at the placement of player offensive and defensive capabilities by season type. To do this, we used the variables season_type, raptor_offense and raptor_defense and visualized the individual placements. 
`season_type`:	Regular season (RS) or playoff (PO)

In order to input the team winning percentage in the following section, we used the following variable, 
`team`:	Basketball-Reference ID of team

In the test for defensive players and WAR during different seasons, we chose to 
use the linear model method to check for correlations of various corresponding variables. 
`raptor_defense`:	Points above average per 100 possessions added by player on defense, using both box and on-off components
`raptor_offense`:	Points above average per 100 possessions added by player on offense, using both box and on-off components


`raptor_total`:	Points above average per 100 possessions added by player on both offense and defense, using both box and on-off components
`war_total`:	Wins Above Replacement between regular season and playoffs


`pace_impact`:	Player impact on team possessions per 48 minutes


### Creating telling variables and cleaning data

```{r variables, echo = FALSE}
nba_players <- nba_players %>% 
  mutate(player_type = case_when(raptor_offense > 0 & raptor_defense > 0 ~ 
                                   "Two-way",
                                raptor_offense > 0 & raptor_defense < 0 ~ 
                                  "Offensive",
                                raptor_offense < 0 & raptor_defense > 0 ~ 
                                  "Defensive",
                                raptor_offense < 0 & raptor_defense < 0 ~ 
                                  "No-way"),
         team_win_pc = case_when(
      team == "MIL" ~ .767,
      team == "TOR" ~ .736,
      team == "BOS" ~ .667,
      team == "IND" ~ .616,
      team == "MIA" ~ .603,
      team == "PHI" ~ .589,
      team == "BRK" ~ .486,
      team == "ORL" ~ .452,
      team == "WAS" ~ .347,
      team == "CHA" ~ .354,
      team == "CHI" ~ .338,
      team == "NYK" ~ .318,
      team == "DET" ~ .303,
      team == "ATL" ~ .299,
      team == "CLE" ~ .292,
      team == "LAL" ~ .732,
      team == "LAC" ~ .681,
      team == "DEN" ~ .630,
      team == "HOU" ~ .611,
      team == "OKC" ~ .611,
      team == "UTA" ~ .611,
      team == "DAL" ~ .573,
      team == "POR" ~ .473,
      team == "MEM" ~ .466,
      team == "PHO" ~ .466,
      team == "SAS" ~ .451,
      team == "SAC" ~ .431,
      team == "NOP" ~ .417,
      team == "MIN" ~ .297,
      team == "GSW" ~ .231)) ## the winning percentage source is cited below
```

Added the following variables to our dataset:

`player_type`: categorizing the player as offensive, defensive etc.

`team_win_pc`: gives each player their associated win percentage. 
Citation: https://www.nba.com/standings

### Interpreting results from visualizations

```{r visualization}
#raptor visualization
ggplot(data = nba_players, aes(x = raptor_defense, y = raptor_offense)) +
  geom_point() + geom_hline(yintercept = 0, lwd = 1, col = "red", lty = 2) +
  geom_vline(xintercept = 0, lwd = 1, col = "red", lty = 2)+facet_wrap(~season_type) +
  labs(title = "Distribution of NBA Players", 
       subtitle = "With the first quadrant being the target area for any player",
       y= "Offensive Raptor Grade", x = "Defensive Raptor Grade")

#all NBA
all_nba <- nba_players %>% 
  filter(player_name == "Giannis Antetokounmpo" |
         player_name == "LeBron James" |
         player_name == "James Harden" |
         player_name == "Anthony Davis" |
         player_name == "Luka Doncic" |
         player_name == "Kawhi Leonard" |
         player_name == "Nikola Jokic" |
         player_name == "Damian Lillard" |
         player_name == "Chris Paul" |
         player_name == "Pascal Siakam" |
         player_name == "Jayson Tatum" |
         player_name == "Jimmy Butler" |
         player_name == "Rudy Gobert" |
         player_name == "Ben Simmons" |
         player_name == "Russell Westbrook") 

all_nba <- all_nba %>% 
  filter(season_type== "RS")

ggplot(data = all_nba) + geom_bar(aes(player_name, raptor_total, 
                                      fill = player_name, xaxt=FALSE), stat = "identity")+theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  labs(title = "Members of the All NBA Teams Total Raptor", 
       y = "Total Raptor Grade",
       fill = "Player Name")

ggplot(data = all_nba) +
  geom_bar(aes(player_name, war_reg_season, fill = player_name, xaxt=FALSE), 
           stat = "identity")+theme(axis.title.x=element_blank(),
                                    axis.text.x=element_blank(), 
                                    axis.ticks.x=element_blank()) +
  labs(title = "Members of the All NBA Teams' WAR",
       fill = "Player Name", y = "WAR")


player_type_data <- nba_players %>%
  filter(player_type != "NA")
  
ggplot(data=player_type_data, aes(x=player_type, y=pace_impact))+geom_violin() +
  labs(title = "Offensive and Two Way Players Play at a Faster Pace",
  subtitle = " Compaired to Defensive and No Way Players",
  y = "Pace of Play Impact", x = "Player Type")


ggplot(data=player_type_data) +
  geom_bar(aes(player_type, raptor_total, fill = player_type), 
             stat = "identity") +
  labs(title = "No Way player types have a terrible Raptor Total",
       subtitle = "While Two Way players have the highest Raptor Total",
       fill = "Player Type",
       x = "Player Type",
       y = "Raptor Total")
 

ggplot(data=player_type_data) +
  geom_bar(aes(player_type, war_total, fill = player_type), 
             stat = "identity") +
  labs(title = "Two Way player types have a great War Total",
       subtitle = "While No Way players have a negative War Total",
       fill = "Player Type",
       x = "Player Type",
       y = "War Total")


```

Explanations:

Graph 1: Graph one shows the distribution of NBA players comparing both their 
offensive and defensive raptor. Each point represents one player and their 
raptors. The first graph shows the players in this years playoffs, and there
seems to be more outliers in it. In both graphs, there looks to be a large 
clump at the axis, meaning that the majority of the NBA has both an offensize 
and defensive raptor around zero. 

Graph 2: Graph two depicts the member of the three all NBA Teams and their 
total raptors. It is interesting to see that only one player was
awarded this honor had a negative total raptor, being Russel Westbrook. Russell
Westbrook was a member of the All NBA Third Team, so if any player
would have negative total raptors it would be him. It is also interesting to 
notice that James Harden had the highest total raptor, because he is a player
known for his incredible offense, but very lackluster defense. I would have 
expected his bad defense to even out his great offensive raptor, but it did not.
His total raptor is just over 5 points higher than the MVP Giannis 
Antetakounmpo.

Graph 3: Graph three shows the WAR of all 15 members of the All-NBA teams. It is
obvious who the league MVP was (Giannis Antetakounmpo) with a WAR well above 
anyone else. What makes this graph interesting is that outside of Antetakounmpo.

Graph 4: Graph three depicts the pace of play by player type. A defensive 
player is a player with a good defensive raptor, a no way player is not
great at either, an offensive player has a good offensive raptor, and a two
way player has a good offensive and defensive raptor. The offensive and two
way players often look to score and play at a fast pace, which this violin
plot clearly shows. The defensive players want to slow the game down and
force the offensive player to be uncomfortable and play slow. The no way
players are often role players, meaning that they do not affect the pace
of the game enough to make it much faster.

Graph 5: While looking into the total raptor for each type of player, we notice
that the no way players have a horrible raptor total. This makes sense, but
visualizing it makes it so much more clear how bad their total raptor really is.
It also effectively shows how important two way players are, as they had the
highest total raptor. In addition, it is interesting to note that defensive
players had a higher total raptor than offensive players.

Graph 6: While looking into the total war for each type of player, we notice
that the two way players have a fantastic war total. This is because ....
In addition, the no way players have a negative war total. The stark comparison
between the two players types is shocking when visualized. Offensive and 
defensive players have a very close total war value, with the defensive players
edging them out.


### 7: Defense

1. Testing if better defensive players have a better WAR in the 
playoffs and regular season.

$H_0$: Better defensive players have the same mean WAR in the regular
season and playoffs as average or below average defenders.

$H_1$: Better defensive players have the better mean WAR in the 
regular season and playoffs than average or below average defenders.

```{r war cor test}

#linear model tests
exp_war_reg <- lm(war_reg_season ~ raptor_defense + raptor_offense, 
              data = nba_players)
tidy(exp_war_reg)

#Conditions test


aug_reg <- augment(exp_war_reg)
aug_reg %>%
  slice(1:3)

#ggplot(data = aug_reg, 
       #aes(x = 1:nrow(nba_players), 
           #y = .resid)) + 
  #geom_point() + 
  #labs(x = "Index", y = "Residual")

ggplot(aug_reg, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(x = "Predicted Price", y = "Residuals")

ggplot(aug_reg, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")

ggplot(aug_reg, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()
```

Expected regular season WAR = 1.70 + 0.142 * Defensive RAPTOR + 0.306 * 
Offensive Raptor

This means that offensive efficiency is actually more important to WAR (wins 
above replacement) than defensive efficiency. The nature of the league has 
definitely changed. Defense used to be the heart and soul of any NBA team, but
now with such high scoring games, the importance of efficient offense seems to
outweigh the importance of efficient defense. 

### 8: Offensive

1. Testing to see whether a player who has a positive offensive raptor has a 
higher mean pace impact.

Method: CLT-based approach

$H_0$: Mean pace_impact when player can play offensive ("Offensive" or "Two-way)
= mean pace_impact when player cannot play offensive but could be "Defensive", 
or "No Way".

$H_1$: Mean pace_impact when player can play offensive 
("Offensive" or "Two-way) > mean pace_impact when player cannot play offensive 
but could be "Defensive", or "No Way".

$/alpha$=0.05

```{r}
nba_players_offense <- nba_players %>%
  mutate(offense=ifelse(raptor_offense > 0, 1, 0))

t.test(pace_impact ~ offense, 
       data = nba_players_offense, 
       var.equal = FALSE,
       alternative = "greater", 
       conf.level = 0.95)
```


Since our P-value of 0.6016 is greater than the alpha level of 0.05, we fail to 
reject the null hypothesis. Since we conducted this test on our population, we 
can say that being able to play offensively does not lend to having a higher
impact on player possessions every 48 minutes. 


### 9: Correlation- RaptorTotal and Winning Percentage

```{r making data setex-2, message = FALSE, warning = FALSE}
team_data_rap <- nba_players %>%
  select(team, team_win_pc, raptor_total) %>%
  group_by(team) %>%
  summarise(avg_rap = mean(raptor_total))

team_win_pct_data <- nba_players %>%
  select(team) %>%
  mutate(team_win_pc = case_when(
    team == "MIL" ~ .767,
    team == "TOR" ~ .736,
    team == "BOS" ~ .667,
    team == "IND" ~ .616,
    team == "MIA" ~ .603,
    team == "PHI" ~ .589,
    team == "BRK" ~ .486,
    team == "ORL" ~ .452,
    team == "WAS" ~ .347,
    team == "CHA" ~ .354,
    team == "CHI" ~ .338,
    team == "NYK" ~ .318,
    team == "DET" ~ .303,
    team == "ATL" ~ .299,
    team == "CLE" ~ .292,
    team == "LAL" ~ .732,
    team == "LAC" ~ .681,
    team == "DEN" ~ .630,
    team == "HOU" ~ .611,
    team == "OKC" ~ .611,
    team == "UTA" ~ .611,
    team == "DAL" ~ .573,
    team == "POR" ~ .473,
    team == "MEM" ~ .466,
    team == "PHO" ~ .466,
    team == "SAS" ~ .451,
    team == "SAC" ~ .431,
    team == "NOP" ~ .417,
    team == "MIN" ~ .297,
    team == "GSW" ~ .231
  )) %>% 
  distinct()

team_data <- inner_join(team_data_rap, team_win_pct_data, by = "team")

ggplot(data = team_data, aes(x = team_win_pc, y= avg_rap)) +geom_point() +
  geom_smooth(method = "lm", se = FALSE, col = "red", lyt = 2, lwd = 1) +
  labs(title = "There are predominantly teams with negative avergae raptor",
       subtitle = "Only five teams that fit the positive raptor, above .500 
       winning percentage", 
       y= "Average Raptor", x = "Team Winning Percantage")

lm_team <- lm(team_win_pc ~ avg_rap, data = team_data)
tidy(lm_team)

aug_reg <- augment(lm_team)
aug_reg %>%
  slice(1:3)

ggplot(aug_reg, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(x = "Predicted Winning Percentage", y = "Residuals")

ggplot(aug_reg, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")

ggplot(aug_reg, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()




```

Expected team winning percentage = 0.52338 + 0.03233(team average raptor)

For every single unit increase in team average raptor, it is expected there will
be an increase in winning percentage of 3.23%.

We tested for equal variance and linearity first. Overall we observed a roughly
linear and equal variance relationship in our data points. Next, we tested for
normality using a QQ plot. This came out less fitting, but we realized there are
only 30 data points so it is expected that the tests won't exactly fit. 

### 10 Correlation - Team Average WAR and Winning Percentage


```{r}

team_data_war <- nba_players %>%
  select(team, team_win_pc, war_reg_season) %>%
  group_by(team) %>%
  summarise(avg_war = mean(war_reg_season))

team_war_data <- inner_join(team_data_war, team_win_pct_data, by = "team")

ggplot(data = team_war_data, aes(x = team_win_pc, y= avg_war)) +geom_point() +
  geom_smooth(method = "lm", se = FALSE, col = "red", lyt = 2, lwd = 1) +
  labs(title = "ADD TITLE",
       subtitle = "ADD SUBTITLE", 
       y= "Average WAR", x = "Team Winning Percantage")

lm_war <- lm(team_win_pc ~ avg_war, data = team_war_data)
tidy(lm_war)

aug_war <- augment(lm_war)
aug_war %>%
  slice(1:3)

ggplot(aug_war, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(x = "Predicted Winning Percentage", y = "Residuals")

ggplot(aug_war, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")

ggplot(aug_war, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()
```

Expected team winning percentage = 0.2012 + 0.3326(team average raptor)

For every single unit increase in team average war, it is expected there will
be an increase in winning percentage of 33.26%. The large increase of win
percentage is due to the fact that the team average Wars are very small and 
very similar, a one point jump would create an outlier, so the 33% increase
to the win percentage is justified.

We tested for equal variance and linearity first. Overall we observed a roughly
linear and equal variance relationship in our data points. Next, we tested for
normality using a QQ plot. This came out less fitting than the raptor test, but 
the normality is good enough to make it accurate.

### 11 Comparing WAR and Raptor Values

```{r compairing War and Raptor Values}

nba_players %>%
 select(war_reg_season, player_name) %>%
  arrange(desc(war_reg_season)) %>%
  slice(1:10)


nba_players %>%
 select(raptor_total, player_name) %>%
  arrange(desc(raptor_total)) %>%
  slice(1:10)
````

This simple look into the players with the top 10 highest Total Raptors and 
Total Wars says a lot about each variable. The top 10 players in terms of War
are household names and allstars. They are some of the best players in the NBA.
Conversely, the top 10 players in terms of Raptor are random players. This being
the case, I would argue that War is a more accurate variable whenever looking 
for the best players. War shows the players that are more important to a teams
win above replacement, meaning that the best players would have the highest 
war. In contrast, the players with the highest raptors could be statistical
anomalies, and the best players clearly do not have the highest raptors. This
is clearest when noticing that Alize Johnson, who average 6.9 minuets per game
for only 17 games last season, has a raptor 100 points over leauge MVP Giannis
Antetokounmpo.



### Discussion
This section is a conclusion and discussion. This will require a summary of what you have learned about your research question along with statistical arguments supporting your conclusions. Also, critique your own methods and provide suggestions for improving your analysis. Issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here. A paragraph on what you would do differently if you were able to start over with the project or what you would do next if you were going to continue work on the project should also be included.
