---
title: "The Recipe for Success in the NBA"
subtitle: "due Friday, Nov 20 at 11:59pm"
author: "Late Round Sleepers: Alex Jackson, Evan Finley, Jack Griffin, Aliya Kamran"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

# 1: Introduction

We are investigating what makes NBA players successful in the context of team 
play. Our reason for choosing this topic of study is that are all NBA fans and eager to 
apply our new knowledge of data science to the game which will allow us to have a 
better understanding of the game and make us more informed viewers. 
Furthermore, basketball statistics are fairly new and it will be interesting to 
discover if our perceived notions of certain players match the statistical data. 

In particular, we define the "success" of NBA players and their teams
based on the following hypothesis tests and questions:

1. Does the average offensive and defensive raptor for players differ between 
the regular season and post season?

2. Does the average raptor differ greatly from the average WAR between members 
of the All NBA teams?

3. How are the variables pace impact, raptor total, and WAR total impacted by 
player type?

4. Do better defensive players have a better WAR in the playoffs than the regular 
season?

5. Do players with better offensive capabilities have greater impact on team 
possessions per 48 minutes as compared to players without offensive 
capabilities?

6. Does the team's average raptor correlate to the team's average winning
percentage?

7. Does the team's average WAR correlate to the team's average winning
percentage?


8. Which statistic, total raptor or total WAR, is better at indicating player 
success? 

Sources:
https://fivethirtyeight.com/features/introducing-raptor-our-new-metric-for-the-modern-nba/


# 2: Data description

```{r environ}
library(tidyverse)
nba_players <- read.csv("https://projects.fivethirtyeight.com/nba-model/2020/latest_RAPTOR_by_team.csv")
```

We are exploring the NBA-Raptor data set produced by Fivethirtyeight. Raptor is 
a player efficiency rating system. This data set has tracked multiple team and 
player analyses over a six year period. Each observation is an NBA player with 
their statistics as variables. It has variables such as players points per game, 
team winning percentage, average team possessions per game, and many more. The 
player data is stats from every NBA games since 2014. The data set has 808 
observations and 23 variables with the investigation focusing on the following 
variables:

`player_name`:	Player name
`season`:	Season
`season_type`:	Regular season (RS) or playoff (PO)
`team`:	Basketball-Reference ID of team
`poss`:	Possessions played
`mp`:	Minutes played
`raptor_offense`:	Points above average per 100 possessions added by player on 
offense, using both box and on-off components
`raptor_defense`:	Points above average per 100 possessions added by player on 
defense, using both box and on-off components
`raptor_total`:	Points above average per 100 possessions added by player on both 
offense and defense, using both box and on-off components
`war_total`:	Wins Above Replacement between regular season and playoffs
`war_reg_season`:	Wins Above Replacement for regular season
`war_playoffs`:	Wins Above Replacement for playoffs
`pace_impact`:	Player impact on team possessions per 48 minutes

# 3: Glimpse of data

```{r glimpse}
glimpse(nba_players)
```

# 4: Methodology

To start with, we looked at the placement of player offensive and defensive 
capabilities by season type. To do this, we used the variables season_type, 
raptor_offense, and raptor_defense and visualized the individual placements.
`season_type`:	Regular season (RS) or playoff (PO)

In order to input the team winning percentage in the following section, we used 
the following variable,
`team`:	Basketball-Reference ID of team

We analyzed the NBA on three different levels, firstly as a whole, secondly by 
player type and thirdly by the All NBA teams of the best players in the game. 
To visualize these analyses we used appropriate graph types such as a scatter 
graph, pie charts, bar graph, box plots as well as a violin graph to 
incorporate density. 

Because the data set consisted of the entire NBA player population, we used 
linear models to look at correlations and the effects of corresponding 
variables on the response. 

In the test for defensive players and WAR during different seasons, we chose to 
use the linear model method to check for correlations of various corresponding 
variables listed below, 
`raptor_defense`:	Points above average per 100 possessions added by player on 
defense, using both box and on-off components
`raptor_offense`:	Points above average per 100 possessions added by player on 
offense, using both box and on-off components
`war_total`:	Wins Above Replacement between regular season and playoffs

Further, we use a CLT based approach to test whether an offensive player has 
more pace in a game, the variables above were used to achieve this. For this 
method we made the assumption that the data was normally distributed.
`pace_impact`:	Player impact on team possessions per 48 minutes

Next, we used a Linear Model in order to examine the relationship between a 
teams average raptor and a teams winning percentage. In doing so, we found the
linear model equation in regards to winning percentage and raptor. 

In the next test, we used a Linear Model to understand the relationship between 
a teams average WAR and their winning percentage. It was a very similar test to 
the example above, and it gave us a different equation in order to calculate 
winning percentage.

Lastly, we simply selected the top 10 players in terms of total raptor and WAR 
and compared them to find possible similarities.


### Creating telling variables and cleaning data


```{r variables, echo = FALSE}
nba_players <- nba_players %>%
  filter(!is.na(war_total))

nba_players <- nba_players %>% 
  mutate(player_type = case_when(raptor_offense > 0 & raptor_defense > 0 ~ 
                                   "Two-way",
                                raptor_offense > 0 & raptor_defense < 0 ~ 
                                  "Offensive",
                                raptor_offense < 0 & raptor_defense > 0 ~ 
                                  "Defensive",
                                raptor_offense < 0 & raptor_defense < 0 ~ 
                                  "No-way"),
         team_win_pc = case_when(
      team == "MIL" ~ .767,
      team == "TOR" ~ .736,
      team == "BOS" ~ .667,
      team == "IND" ~ .616,
      team == "MIA" ~ .603,
      team == "PHI" ~ .589,
      team == "BRK" ~ .486,
      team == "ORL" ~ .452,
      team == "WAS" ~ .347,
      team == "CHA" ~ .354,
      team == "CHI" ~ .338,
      team == "NYK" ~ .318,
      team == "DET" ~ .303,
      team == "ATL" ~ .299,
      team == "CLE" ~ .292,
      team == "LAL" ~ .732,
      team == "LAC" ~ .681,
      team == "DEN" ~ .630,
      team == "HOU" ~ .611,
      team == "OKC" ~ .611,
      team == "UTA" ~ .611,
      team == "DAL" ~ .573,
      team == "POR" ~ .473,
      team == "MEM" ~ .466,
      team == "PHO" ~ .466,
      team == "SAS" ~ .451,
      team == "SAC" ~ .431,
      team == "NOP" ~ .417,
      team == "MIN" ~ .297,
      team == "GSW" ~ .231)) ## the winning percentage source is cited below
```

Added the following variables to our dataset:

`player_type`: categorizing the player as offensive, defensive, etc.

`team_win_pc`: gives each player their associated win percentage. 
Citation: https://www.nba.com/standings


# 5: Testing, results and visualizations


```{r visualization}
#raptor visualization
ggplot(data = nba_players, aes(x = raptor_defense, y = raptor_offense)) +
  geom_point() + geom_hline(yintercept = 0, lwd = 1, col = "red", lty = 2) +
  geom_vline(xintercept = 0, lwd = 1, col = "red", lty = 2)+facet_wrap(~season_type) +
  labs(title = "G1: Distribution of NBA Players", 
       subtitle = "With the first quadrant being the target area for any player",
       y= "Offensive Raptor Grade", x = "Defensive Raptor Grade")

#all NBA
all_nba <- nba_players %>% 
  filter(player_name == "Giannis Antetokounmpo" |
         player_name == "LeBron James" |
         player_name == "James Harden" |
         player_name == "Anthony Davis" |
         player_name == "Luka Doncic" |
         player_name == "Kawhi Leonard" |
         player_name == "Nikola Jokic" |
         player_name == "Damian Lillard" |
         player_name == "Chris Paul" |
         player_name == "Pascal Siakam" |
         player_name == "Jayson Tatum" |
         player_name == "Jimmy Butler" |
         player_name == "Rudy Gobert" |
         player_name == "Ben Simmons" |
         player_name == "Russell Westbrook") 

all_nba <- all_nba %>% 
  filter(season_type== "RS")

ggplot(data = all_nba, aes(x=factor(1), fill=player_type))+
  geom_bar(width = 1)+
  coord_polar("y") +
  labs(title =  "G2: Two-Way players have the highest representation 
  on the All-NBA teams",
       fill = "Player Type")+
  theme(axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())

```

```{r pie chart fix}
ggplot(data = nba_players, aes(x=factor(1), fill=player_type))+
  geom_bar(width = 1)+
  coord_polar("y") +
  labs(title = 
         "G3: Defensive and No-way Players make up the majority of the NBA",
       fill = "Player Type")+
  theme(axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank())  

```

```{r vis. cont.}
ggplot(data = all_nba) + geom_bar(aes(player_name, raptor_total, 
                                      fill = player_name, xaxt=FALSE), 
                                  stat = "identity")+
  theme(axis.title.x=element_blank(), 
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank()) +
  labs(title = "G4: Members of the All NBA Teams Total Raptor", 
       y = "Total Raptor Grade",
       fill = "Player Name")

ggplot(data = all_nba) +
  geom_bar(aes(player_name, war_reg_season, fill = player_name, xaxt=FALSE), 
           stat = "identity")+theme(axis.title.x=element_blank(),
                                    axis.text.x=element_blank(), 
                                    axis.ticks.x=element_blank()) +
  labs(title = "G5: Members of the All NBA Teams' WAR",
       fill = "Player Name", y = "WAR")


player_type_data <- nba_players %>%
  filter(season_type == "RS", !is.na(war_reg_season))
  
ggplot(data=player_type_data, aes(x=player_type, y=pace_impact))+
  geom_violin() +
  labs(title = "G6: Offensive and Two Way Players Play at a Faster Pace",
  subtitle = " Compaired to Defensive and No Way Players",
  y = "Pace of Play Impact", x = "Player Type")


ggplot(data=player_type_data, aes(x=player_type, y=raptor_total)) +
  geom_boxplot(aes(color = player_type))+
  labs(title = "G7: No Way player types have a terrible Raptor Total",
       subtitle = "While Two Way players have the highest Raptor Total",
       color = "Player Type",
       x = "Player Type",
       y = "Raptor Total")
 

ggplot(data=player_type_data, aes(x=player_type, y=war_total)) +
  geom_boxplot(aes(color = player_type)) +
  labs(title = "G8: Two Way player types have a great War Total",
       subtitle = "While No Way players have a negative War Total",
       color = "Player Type",
       x = "Player Type",
       y = "War Total")

```

Explanations:

Graph 1: Graph one shows the distribution of NBA players, comparing both their 
offensive and defensive raptor. Each point represents one player and their 
raptors. The first graph shows the players in this year's playoffs, and there
seems to be more outliers in it. In both graphs, there looks to be a large 
clump at the axis, meaning that the majority of the NBA has both an offensive 
and defensive raptor around zero. 

Graphs 2 and 3: In the second and third graphs, we look at the breakdown of
player types by the All-NBA teams and the entire NBA. The first pie chart looks
at the 15 All-NBA players, where Two-Way players dominate (8/15). In comparison,
the pie chart of the entire NBA shows that Two-Way players nearly have the
smallest proportional representation. What this signifies is that Two-way players
are a more "elite" breed and find more success in the NBA being elite a both
ends of the floor. Defensive players have a small representation of All-NBA
players, yet the most of any type in the NBA, so that signifies the difficulty
of being the best in the league as a Defensive-type player. 

Graph 4: Graph two depicts the member of the three All NBA Teams and their 
total raptors. It is interesting to see that only one player who was
awarded this honor had a negative total raptor, Russel Westbrook. Russell
Westbrook was a member of the All NBA Third Team, so if any All NBA player
would have negative total raptor it makes sense that it is him. It is also 
interesting to notice that James Harden had the highest total raptor, because 
he is a player known for his incredible offense, but very lackluster defense. 
I would have expected his bad defense to even out his great offensive raptor, 
but it did not, which speaks to the changing nature of the league, which values
offense. His total raptor is just over 5 points higher than the MVP Giannis 
Antetakounmpo.

Graph 5: Graph three shows the WAR of all 15 members of the All-NBA teams. It is
not obvious who the league MVP was (Giannis Antetakounmpo), due to the fact
that James Harden is towering above everyone else. Harden finished 3rd in
league MVP voting, which is peculiar because based on this metric, he would
have been by far the most valuable to lose. Antetakounmpo on the other hand,
barely reaches the title of second highest and could easily be mixed up among
other All-NBA players. This, combined with Graph 2, raises the question of why
Antetakounmpo was MVP instead of Harden?


Graph 6: Graph three depicts the pace of play by player type. A defensive 
player is a player with a good defensive raptor, an offensive player has a 
good offensive raptor, a no way player is not great at either, and a two
way player has a good offensive and defensive raptor. The offensive and two
way players often look to score and play at a fast pace, which this violin
plot clearly shows due to the larger influx of those players above the 0.0 line. 
The defensive players want to slow the game down and force the offensive player 
to be uncomfortable and play slowly. The no way players are often role players, 
meaning that they do not affect the pace of the game enough to make it much faster.
There are a few No-way players and Two-way players that greatly separate 
themselves from their respective trends. This can be explained as outliers who
played few minutes over the course of the season.

Graph 7: When looking at the total raptor for each type of player, we notice
that the no way players have a incredibly low raptor total. This makes sense, but
visualizing it makes it much more clear how low their total raptor really is.
It also effectively shows how important two way players are, as they had the
highest total raptor. In addition, it is interesting to note that defensive
players had a higher total raptor than offensive players.

Graph 8: While looking into the total war for each type of player, we notice
that the two way players have a wildly high war total. This is because WAR takes
into account a player's effect on the team on both offense and defense, so two-
way players generate higher WARs typically becuase they are excel on both ends
of the floor. In addition, the no way players have a negative war total because
their impact on the game actually hurts their teams.The stark comparison between 
the two players types is shocking when visualized. Offensive and defensive players 
have a very close total war value, with defensive players slightly ahead.


### Which is more effective on WAR, offense or defense?

```{r war-cor-test}

#linear model tests
exp_war_reg <- lm(war_reg_season ~ raptor_defense + raptor_offense, 
              data = nba_players)
tidy(exp_war_reg)

#Conditions test

aug_reg <- augment(exp_war_reg)
aug_reg %>%
  slice(1:3)

ggplot(aug_reg, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(tile = "Equal variance and linearity test",
       x = "Predicted WAR", y = "Residuals")

ggplot(aug_reg, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")+
  labs(title= "Normality test")

ggplot(aug_reg, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line() +
  labs(title = "Normality test")

## R^2 value
glance(exp_war_reg)$r.squared

nba_players %>% 
  summarise(cor_offense = cor(war_reg_season, raptor_offense))

nba_players %>% 
  summarise(cor_defense = cor(war_reg_season, raptor_defense))

```
We would love to look at the relationship between a players WAR and their
offensive and defensive raptor grades, but they are just not that connected.
In the linear model test we conducted, the results were as follows:

Expected regular season WAR = 1.02 + 0.026 * Defensive RAPTOR + 0.125 * 
Offensive Raptor

This means that offensive efficiency is actually more important to WAR (wins 
above replacement) than defensive efficiency. The nature of the league has 
definitely changed. Defense used to be the most critical element of winning a 
game, but now with such high scoring games, the importance of efficient offense 
seems to outweigh the importance of efficient defense.

This analyses above isn't entirely wrong. After performing two correlation tests,
we found that offensive raptor was more highly correlated with WAR than defesive
raptor. The issue is that our tests for linearity, equal variance and normality
all came back very concerning. Additionally, the R-squared showed that roughly 
only 9.4% of the variability in WAR of players can be explained by raptor
offense and raptor defense for the regular season. This reveals that despite the
linear model producing believable results, it's not entirely predictive and should
be used with caution. 



### Offensive

1. Testing to see whether a player who has a positive offensive raptor has a 
higher mean pace impact.

Method: CLT-based approach

$H_0$: Mean pace_impact when player has offensive capabilities = mean 
pace_impact when player cannot play offensively. 

$H_1$: Mean pace_impact when player has offensive capabilities is greater than 
mean pace_impact when player cannot play offensively.

$/alpha$=0.05

```{r}
nba_players_offense <- nba_players %>%
  mutate(offense=ifelse(raptor_offense > 0, 1, 0))

t.test(pace_impact ~ offense, 
       data = nba_players_offense, 
       var.equal = FALSE,
       alternative = "greater", 
       conf.level = 0.95)
```

Since our P-value of 0.6016 is greater than the alpha level of 0.05, we fail to 
reject the null hypothesis. Since we conducted this test on the total 
population of all NBA players, we can say that being able to play offensively 
does not lend to having a higher impact on player possessions every 48 minutes. 

The t-test is a tool to test as assumption for the population, and since the 
data set already contains the population, we can say with greater certainty 
that the above implication holds. Since the population is constantly changing 
in the NBA, it would still be valuable to look at this test despite the 
limitations. 

### Correlation- RaptorTotal and Winning Percentage

```{r making data setex-2, message = FALSE, warning = FALSE}
team_data_rap <- nba_players %>%
  select(team, team_win_pc, raptor_total) %>%
  group_by(team) %>%
  summarise(avg_rap = mean(raptor_total))

team_win_pct_data <- nba_players %>%
  select(team) %>%
  mutate(team_win_pc = case_when(
    team == "MIL" ~ .767,
    team == "TOR" ~ .736,
    team == "BOS" ~ .667,
    team == "IND" ~ .616,
    team == "MIA" ~ .603,
    team == "PHI" ~ .589,
    team == "BRK" ~ .486,
    team == "ORL" ~ .452,
    team == "WAS" ~ .347,
    team == "CHA" ~ .354,
    team == "CHI" ~ .338,
    team == "NYK" ~ .318,
    team == "DET" ~ .303,
    team == "ATL" ~ .299,
    team == "CLE" ~ .292,
    team == "LAL" ~ .732,
    team == "LAC" ~ .681,
    team == "DEN" ~ .630,
    team == "HOU" ~ .611,
    team == "OKC" ~ .611,
    team == "UTA" ~ .611,
    team == "DAL" ~ .573,
    team == "POR" ~ .473,
    team == "MEM" ~ .466,
    team == "PHO" ~ .466,
    team == "SAS" ~ .451,
    team == "SAC" ~ .431,
    team == "NOP" ~ .417,
    team == "MIN" ~ .297,
    team == "GSW" ~ .231
  )) %>% 
  distinct()

team_data <- inner_join(team_data_rap, team_win_pct_data, by = "team")

ggplot(data = team_data, aes(x = team_win_pc, y= avg_rap)) +geom_point() +
  geom_smooth(method = "lm", se = FALSE, col = "red", lyt = 2, lwd = 1) +
  labs(title = "There are predominantly teams with negative avergae raptor",
       subtitle = "Only five teams that fit the positive raptor, above .500 
       winning percentage", 
       y= "Average Raptor", x = "Team Winning Percantage")

lm_team <- lm(team_win_pc ~ avg_rap, data = team_data)
tidy(lm_team)

aug_reg <- augment(lm_team)
aug_reg %>%
  slice(1:3)

ggplot(aug_reg, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(title = "Equal variance and Linearity test", 
       x = "Predicted Winning Percentage", y = "Residuals")

ggplot(aug_reg, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(title = "Normality test", x = "Residuals")

ggplot(aug_reg, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()+
  labs(title = "Normality test")




```

Expected team winning percentage = 0.52338 + 0.03233(team average raptor)

For every single unit increase in team average raptor, it is expected there will
be an increase in winning percentage of 3.23%.

We tested for equal variance and linearity first. Overall we observed a roughly
linear and equal variance relationship in our data points. Next, we tested for
normality using a QQ plot. This came out less fitting, but we realized there are
only 30 data points so it is expected that the tests won't exactly fit. 

### Correlation - Team Average WAR and Winning Percentage


```{r}

team_data_war <- nba_players %>%
  select(team, team_win_pc, war_reg_season) %>%
  group_by(team) %>%
  summarise(avg_war = mean(war_reg_season))

team_war_data <- inner_join(team_data_war, team_win_pct_data, by = "team")

ggplot(data = team_war_data, aes(x = team_win_pc, y= avg_war)) +geom_point() +
  geom_smooth(method = "lm", se = FALSE, col = "red", lyt = 2, lwd = 1) +
  labs(title = "A higher average team WAR leads to a higher winning percentage",
       subtitle = "(For the most part)", 
       y= "Average WAR", x = "Team Winning Percentage")

lm_war <- lm(team_win_pc ~ avg_war, data = team_war_data)
tidy(lm_war)

aug_war <- augment(lm_war)
aug_war %>%
  slice(1:3)

ggplot(aug_war, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(title = "Equal variance and linearity test",
       x = "Predicted Winning Percentage", y = "Residuals")

ggplot(aug_war, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(title= "normality test",
      x = "Residuals")

ggplot(aug_war, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()+
  labs(title = "normality test")
```

Expected team winning percentage = 0.2012 + 0.3326(team average raptor)

For every single unit increase in team average WAR, it is expected there will
be an increase in winning percentage of 33.26%. The large increase of win
percentage is due to the fact that the team average WARs are very small and 
very similar, a one point jump would create an outlier, so the 33% increase
to the win percentage is understandable.

We tested for equal variance and linearity first. Overall we observed a roughly
linear and equal variance relationship in our data points. Next, we tested for
normality using a QQ plot. This came out less fitting than the raptor test, but 
the normality is close enough to pass the test. With only 30 datapoints, it is
hard to tell an exact fit from these tests, so the overall linear model should
be used with slight caution.

### Comparing WAR and Raptor Values

```{r compairing War and Raptor Values}

nba_players %>%
 select(war_reg_season, player_name) %>%
  arrange(desc(war_reg_season)) %>%
  slice(1:10)


nba_players %>%
 select(raptor_total, player_name) %>%
  arrange(desc(raptor_total)) %>%
  slice(1:10)
````

This simple test looks at the players with the top 10 highest Total Raptors and 
Total WARs and says a lot about each variable. The top 10 players in terms of WAR
are all household names and allstars. They are some of the best players in the NBA.
Conversely, the top 10 players in terms of Raptor are random players. This being
the case, I would argue that WAR is a more accurate variable whenever looking 
for the best players. WAR shows the players that are more important to a team's
win above replacement, meaning that the best players would have the highest 
WAR. In contrast, the players with the highest raptors could be statistical
anomalies, and the best players clearly do not have the highest raptors. This
is clearest when noticing that Alize Johnson, who average 6.9 minuets per game
for only 17 games last season, has a raptor 100 points higher than league MVP 
Giannis Antetokounmpo.

# 6: Discussion

Before we analyzed this dataset, we assumed that defensive players were most 
valuable to a team. However, based on our data, we have come to the conclusion 
that offensive players are actually more important. Moreover, two way players 
are invaluable for a team, and are key to a teams overall success. Offensive 
raptor correlates more highly to WAR than defensive raptor. We associate WAR 
with individual success, as it is a variable that shows a players overall 
importance to the team in terms of wins above replacement. In addition, when 
looking to find the more valuable players, using the WAR variable is more 
useful, as there are statistical outliers in the raptor data that do not always 
portray complete accuracy. We also learned a lot in our visualizations, such as 
the importance and lack there of two way players. In addition, we learned that 
the majority of the players in the NBA had an average raptor around zero, making 
there a lot of average players in the NBA.  

In terms of suggestions for improving our analysis, we could have made sure that 
our correlation tests worked better. Especially in our first correlation test, 
the QQ plots showed some non-normality, and we could have come up with better 
conclusions if our QQ plots were more accurate. In addition, we performed a 
T-test on a full population. Due to the fact that NBA changes every year, 
performing a T-test on a full population could suggest accurate predictions 
about future years. In terms of our data, most of the variables are numeric 
based, and do not necessarily contain all of the nuances of the NBA, such as 
coaching, team morale, synergy between teammates, and injuries.

If we were to start over, we would have chosen a similar data set but one with 
a random population of NBA players throughout history. If we had done this, we 
would have been able to perform sample population tests. In addition, we would 
have chosen a data set with a few simpler statistics, such as points per game, 
rebounds per game, and assists per game, in order to analyze the change of these 
stats over the years. The NBA has become much more an offensive league, and it 
would have been interesting to observe that change analytically. If we were to 
continue to work on this project, we would try to find linear models that fit 
effectively, or we would have continued to try to make the data fit (we did not 
fully learn how to transform complex variables).




