HI
---
title: "Proposal"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(tidyverse)
nba_players <- read.csv("https://projects.fivethirtyeight.com/nba-model/2020/latest_RAPTOR_by_team.csv")
```
## R Markdown


###Section 1 - Introduction


We are investigating what makes NBA players successful in various seasons. We
are all NBA fans and looking into statistics about the NBA will allow us to 
have a better understanding for the game and make us more informative fans. 
Basketball statistics are fairly new and it will be interesting to discover if
our perceived notions of certain players match the statistical data. We want
to look at if defensive or offensive efficiency affects winning percentage the
most.
In particular, we want to test the following hypothesis

Better defensive players have a better winning percentage in the playoffs and 
regular season.

Players who score the most do not have the highest winning percentage.

Players who have the most offensive possessions win more games.

Sources:
https://fivethirtyeight.com/features/introducing-raptor-our-new-metric-for-the-modern-nba/


###Section 2 - Data description


We wish to explore the NBA-Raptor data set. Raptor is a player efficiency rating 
system.This data set has tracked multiple team and player analysis over a six 
year period. Each observation is an NBA player with variables about their 
statistics. It has variables such as players points per game, team winning 
percentage, average team possessions per game, and many more. The player data
is stats from every NBA games since 2014. Raptor was made by Fivethirtyeight. 



###Section 3 - Glimpse of data

```{r section-3}
glimpse(nba_players)
```

### Methodology

### Results

### Adding Variables
```{r variables}
nba_players <- nba_players %>% 
  filter(war_reg_season != 0)
  

mean_war <- nba_players %>% 
  summarise(mean_war = mean(war_reg_season))
mean_war

nba_players <- nba_players %>% 
  mutate(war_reg_good = case_when(war_reg_season > mean_war ~ TRUE,
                                  war_reg_season <= mean_war ~ FALSE),
         player_type = case_when(raptor_offense > 0 & raptor_defense > 0 ~ 
                                   "Two-way",
                                raptor_offense > 0 & raptor_defense < 0 ~ 
                                  "Offensive",
                                raptor_offense < 0 & raptor_defense > 0 ~ 
                                  "Defensive",
                                raptor_offense < 0 & raptor_defense < 0 ~ 
                                  "No-way"))
```



```{r team win pc}
nba_players <- nba_players %>%
  mutate(team_win_pc = case_when(
    team == "MIL" ~ .767,
    team == "TOR" ~ .736,
    team == "BOS" ~ .667,
    team == "IND" ~ .616,
    team == "MIA" ~ .603,
    team == "PHI" ~ .589,
    team == "BRK" ~ .486,
    team == "ORL" ~ .452,
    team == "WAS" ~ .347,
    team == "CHA" ~ .354,
    team == "CHI" ~ .338,
    team == "NYK" ~ .318,
    team == "DET" ~ .303,
    team == "ATL" ~ .299,
    team == "CLE" ~ .292,
    team == "LAL" ~ .732,
    team == "LAC" ~ .681,
    team == "DEN" ~ .630,
    team == "HOU" ~ .611,
    team == "OKC" ~ .611,
    team == "UTA" ~ .611,
    team == "DAL" ~ .573,
    team == "POR" ~ .473,
    team == "MEM" ~ .466,
    team == "PHO" ~ .466,
    team == "SAS" ~ .451,
    team == "SAC" ~ .431,
    team == "NOP" ~ .417,
    team == "MIN" ~ .297,
    team == "GSW" ~ .231
  ))
           
```





1. Testing if better defensive players have a better WAR in the 
playoffs and regular season.

$H_0$: Better defensive players have the same WAR in the regular
season and playoffs as average or below average defenders.

$H_1$: Better defensive players have the better WAR in the 
regular season and playoffs than average or below average defenders.

```{r war cor test}
nba_players <- nba_players %>% 
  mutate(log_war_reg = log(war_reg_season)) 

#checking mean WAR
mean_war <- nba_players %>% 
  summarise(mean_war = mean(log_war_reg))
mean_war

mean_war_po <- playoff_players %>% 
  summarise(mean_war = mean(war_playoffs))
mean_war_po

ggplot(data = nba_players, aes(x = raptor_defense, y = raptor_offense)) +
  geom_point() + geom_hline(yintercept = 0, lwd = 1, col = "red", lty = 2) +
  geom_vline(xintercept = 0, lwd = 1, col = "red", lty = 2)


#linear model tests
exp_war_reg <- lm(log_war_reg ~ raptor_defense + raptor_offense, 
              data = nba_players)
tidy(exp_war_reg)

#Conditions test


aug_reg <- augment(exp_war_reg)
aug_reg %>%
  slice(1:3)

ggplot(data = aug_reg, 
       aes(x = 1:nrow(nba_players), 
           y = .resid)) + 
  geom_point() + 
  labs(x = "Index", y = "Residual")

ggplot(aug_reg, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() + 
  geom_hline(yintercept = 0, lwd = 2, col = "red", lty = 2) +
  labs(x = "Predicted Price", y = "Residuals")

ggplot(aug_reg, mapping = aes(x = .resid)) +
  geom_histogram(bins = 15) + 
  labs(x = "Residuals")

ggplot(aug_reg, mapping = aes(sample = .resid)) +
  stat_qq() + 
  stat_qq_line()

exp_war_po <- lm(war_playoffs ~ raptor_defense + raptor_offense, 
              data = playoff_players)
tidy(exp_war_po)
```

Expected regular season WAR = 1.70 + 0.142 * Defensive RAPTOR + 0.306 * 
Offensive Raptor

Expected playoff WAR = 0.470 + 0.104 * Defensive RAPTOR + 0.152 * Offensive 
Raptor

This means that offensive efficiency is actually more important to WAR (wins 
above replacement) than defensive efficiency in both the regular season and 
playoffs.


2. Do the players on the team's raptor total correlate to the teams winning
percentage?

```{r making data setex-2}
data_for_ex_2 <- nba_players %>%
  select(player_name, team, raptor_total, team_win_pc) 
 

```

```{r ex-2}
mil_data <- data_for_ex_2 %>%
  filter(team == "MIL") %>%
  mutate(team_average_total_raptor = case_when(
    team == "MIL" ~ mean(raptor_total)
  ))

raptors_data <- data_for_ex_2 %>%
  filter(team == "TOR") %>%
  mutate(team_average_total_raptor = case_when(
    team == "TOR" ~ mean(raptor_total)
  ))

celtics_data <- data_for_ex_2 %>%
  filter(team == "BOS") %>%
  mutate(team_average_total_raptor = case_when(
    team == "BOS" ~ mean(raptor_total)
  ))

pacers_data <- data_for_ex_2 %>%
  filter(team == "IND") %>%
  mutate(team_average_total_raptor = case_when(
    team == "IND" ~ mean(raptor_total)
  ))

lakers_data <- data_for_ex_2 %>%
  filter(team == "LAL") %>%
  mutate(team_average_total_raptor = case_when(
    team == "LAL" ~ mean(raptor_total)
  ))

clippers_data <- data_for_ex_2 %>%
  filter(team == "LAC") %>%
  mutate(team_average_total_raptor = case_when(
    team == "LAC" ~ mean(raptor_total)
  ))

nuggets_data <- data_for_ex_2 %>%
  filter(team == "DEN") %>%
  mutate(team_average_total_raptor = case_when(
    team == "DEN" ~ mean(raptor_total)
  ))

rockets_data <- data_for_ex_2 %>%
  filter(team == "HOU") %>%
  mutate(team_average_total_raptor = case_when(
    team == "HOU" ~ mean(raptor_total)
  ))

```

```{r combining data sets}
total_data_ex_2 <-rbind(rockets_data, nuggets_data, clippers_data, lakers_data,
                        pacers_data, celtics_data, raptors_data, mil_data)

```

```{r correlation between team win pct and team average raptor}
total_data_ex_2 %>%
  summarise(cor_team_win_team_raptor = cor(team_win_pc, team_average_total_raptor))


```
There is a weak correlation between team total raptor and team winning percentage.